Theory about Normal Basis on $F(2^k)$

### 2016-12-11 dissertation/normal.tex
dissertation v3:milestone

\chapter{Functional Verification of Sequential Normal Basis Multiplier}
\label{ch:normal}
In order to utilize our traversal algorithm, it is necessary to find out
a sort of suitable circuit benchmarks which is easy to compute its
Gr\"obner basis (GB). From the work of {\it Lv} \cite{lv:phd},
we learn that arithmetic circuits in Galois field (GF) is
convertible to an ideal of circuit polynomials, and the 
ideal generators form a GB themselves when applying reverse topological
term order. Furthermore, according to the work of {\it Pruss et al.}
\cite{pruss:tcad15}, with a limited computation complexity,
we can abstract the word-level signature of an arithmetic 
component working in GF. Thus, we consider the possibility 
of applying our traversal algorithm on sequential Galois
field circuits. In each frame, we can use the techniques 
from \cite{pruss:tcad15} to abstract the word-level
signature of the combinational logic, which corresponds
to the transition function in our traversal algorithm.
As a result, we manage to find a type of sequential GF multiplier
which we can apply our traversal algorithm to actually 
verify its functional correctness.

\section{Motivation}
\label{sec:normal_motiv}
From the preliminaries about FSMs in Section \ref{sec:reach}, we learn that the
Moore machine does not rely on inputs for state transitions. 
As depicted in Figure \ref{fig:Moore}(a), a typical Moore machine implementation
consists of combinational logic component and register files, where
$r_0,\dots,r_k$ are present state (PS) variables 
standing for state inputs (SI), and $r_0',\dots,r_k'$ are next state (NS) variables standing for
state outputs (SO). Figure \ref{fig:Moore}(b) shows the state transition graph (STG) of 
a Moore machine with $k+1$ distinct states. We notice that it forms a simple chain,
with $k$ consecutive transitions the machine reaches final state $R_k$.

@@

In practice,
some arithmetic components are designed in sequential circuits similar to the structure in 
Figure \ref{fig:Moore}(a). Initially the operands are loaded into the registers, 
then the closed circuit executes without taking any additional information from outside,
and store the results in registers after $k$ clock cycles. Its behavior can be described using
STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, $R_{init}$ is the initial
state (usually reset to all zeros), $R_1$ to $R_{k-1}$ are intermediate results stored as SO of current state and SI
for next state, and $R_k$ (or $R_{final}$) is the final result given by arithmetic circuits (and equals to the answer
to arithmetic function when circuit is working functional correctly).
This kind of design results in 
reusing a smaller combinational logic component such that the area cost is greatly optimized.
An example of these designs is the sequential GF arithmetic multipliers described in Chapter \ref{ch:prelim_GF}.
However, it also brings difficulties in verifying the the circuit functions.

@@

Conventional methods to such a sequential circuit may consist of unrolling the circuit for 
$k$ time-frames, and performing an equivalence checking between the unrolled machine and
the specification function. However, the number of gates will grow fast when doing unrolling
on bit-level. Meanwhile the structural similarity based equivalence checking techniques 
will fail when the sequential circuit is highly customized and optimized from the naive specification 
function. As a result, conventional techniques is grossly inefficient for large circuits.
Therefore, a new method based on our proposed word-level FSM traversal technique is worthy to be explored.

\section{Formal Verification of Normal Basis Multiplier using Gr\"obner Basis}
The gate-level design of a NB multiplier can be generated using appoaches introduced in 
Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
an approach similar to that in Chapter \ref{ch:reacha}. 
First we introduce the sketch of our approach using abstraction term order (ATO)
mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
the concept ``RATO", which is previously used in Section \ref{sec:improve}.
% The following part borrows contents
% from my own conference paper \cite{myDATE}.
\subsection{Implicit Unrolling based on Abstraction with ATO}
\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
\caption{Architecture of a combinational GF multiplier}
\label{fig:comb_RAB}}
\end{figure}

The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
$intermediate~variables~>~R~>~A,B$ for the circuit in Figure \ref{fig:comb_RAB} to compute Gr\"obner basis, 
the function of the combinational logic component can be abstracted as 
$$R = \Func(A,B)$$
To verify the functional correctness of a combinational NB multiplier (e.g. Mastrovito multiplier
or Montgomery multiplier), the function given by abstraction will be computed as
$$R=A\cdot B$$
While in the sequential case, the function of combinational logic only fulfills a part of the multiplication.
For example, in the RH-SMPO introduced in Section \ref{sec:nbdesign}, 
the combinational logic actually implements  $F_{k-1}$ in Equation \ref{eqn:aux},
while computing all of $F_i$ requires $k$-cycle unrolling of the circuit.
Nevertheless, the abstraction still provides a word-level 
representation which can be more efficient for unrolling than bit-level expressions.
In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
explicit unrolling and avoid bit-blasting problem.

For 2-input sequential NB multipliers, abstraction is utilized to implement following algorithm:
%\vspace{-0.1in}
\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
   state ideal for output $R (=0)$, inputs $\mathcal{G}(A_{init}), \mathcal{H}(B_{init})$} 
\KwOut{Circuit function $R_{final}$ after $k$ cycles of $A_{init}, B_{init}$}
  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  \tcc{$A, B$ are word-level variables, solutions to polynomial equations 
  $\mathcal{G},\mathcal{H}$ denote the initial values of preloaded operands $A_{init},B_{init}$}
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	\tcc{Using projections of varieties from abstraction theory, we get NS variables $R',A',B'$
	in terms of PS $A,B$}
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
	\tcc{By replacing NS variables with PS variables we push it to next time-frame}
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
\DecMargin{1em}

@@

We follow  the sequential GF circuit model of
Figure \ref{fig:sequential}, with word-level variables $A, B, R$
denoting the {\it present states (PS)} and $A', B', R'$ denoting the {\it next
  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
store the result, and $A, B\ (A', B')$ store input operands. E.g.,
for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
0$) are the initial values (operands) loaded into the registers,  and
$R = \Func(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
result after $k$-cycles. Our approach aims to find this polynomial
representation for $R$.

Each gate in the combinational logic is represented by a Boolean
polynomial. To 
this set of Boolean polynomials, we append the polynomials that define
the word-level to bit-level relations for PS and NS variables ($A =
\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
as ideal $J = \langle 
f_1, \dots, f_s \rangle \subset \Fkk[x_1, \dots, x_d, R, R', A, A', B,
  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
  of the circuit. The ideal of vanishing polynomials $J_0$ is also included, and
then the implicit FSM unrolling problem is setup for abstraction.

The configurations of the flip-flops are the states of the
machine. Since the set of states is a finite set of points, we
can consider it as the variety of an ideal related to the circuit.
Moreover, since we are interested in
the function encoded by the state variables (over $k$-time
frames), we can project this variety on the word-level state
variables, starting from the initial state $A_{init}, B_{init}$.
Projection of varieties (geometry) corresponds to elimination ideals
(algebra), and can be analyzed via \Grobner bases. Therefore, we
employ a \Grobner basis computation with ATO: we use a {\it lex term
  order} with {\it bit-level variables} 
$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
allows to eliminate all the bit-level variables 
%(corresponding to the combinational logic and the state variables),
%so as to 
and derives a representation only in terms of words. 
Consequently, $k$-successive \Grobner basis computations implicitly
unroll the machine, and provide word-level algebraic $k$-cycle
abstraction for $R'$ as $R' = \Func(A_{init}, B_{init})$.

@@

In the practice of sequential GF multiplier verification, the combinational 
logic actually implements function not only related to current operands $A$ and $B$,
but also involved with PS variable (i.e. temporary product) $R$. Which can be 
obtained using the abstraction:
$$R' = \Func(A,B,R)$$

@@

\begin{Example}[Functional verification of 5-bit RH-SMPO]
\label{ex:RHSMPO}
Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO. The transition function for
operands $A,B$ is doing cyclic shift, while transition function for $R$ has to be computed through Gr\"obner basis
abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
all gates in combinational logic block and definition of word-level variables.
$$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$

@@

According to line 5 of Algorithm \ref{alg:modified}, we merge $J_{ckt}$, $J_0$ and $from^i$, then compute its
Gr\"obner basis with abstraction term order. 
Concretely, in the first iteration $from^0$ contains three generators. The first one describes PS variable $R$ --
temporary (or initial, in first iteration) product, which equals to 0 according to the mechanism of sequential 
GF multiplier. The polynomial is written as 
$$R$$
The other two polynomials describes PS variable $A,B$ -- current multiplication operands, and we write them 
in parameters $A_{init}, B_{init}$. Since this is the first iteration, they have the same form with $f_{16}$ and $f_{17}$:
\begin{align*}
& A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18} \\
& B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}
\end{align*}
After computing the Gr\"obner basis of $J_{ckt}+J_0+from^0$ using ATO 
$$\text{all other bit-level variables} > R > \{R',A_{init},B_{init}\}$$
there is a polynomial in form of $R'+\Func(A_{init},B_{init})$,
which should be included by $to^{i+1}$. $to^{i+1}$ also exclude next state variable $A'$ and $B'$, instead we 
redefine $A_{init}$ and $B_{init}$ using next state bit-level variables $\{a_i', b_j'\}$. Next state Bit-level variables
$a_i' = a_{i-1\pmod k}, b_j' = b_{j-1\pmod k}$ according to definition of cyclic shift.

@@

\subsection{Overcome Computational Complexity using RATO}
% although commented, but I insist refer to previous chapter
% TODO experiment on naive slimgb on SMPO
We implement Algorithm \ref{alg:modified} in \textsc{Singular}, with the same setup and environment 
with the experiment in Section \ref{sec:exp_reacha}. The result on verifying Agnew's SMPO is shown 
in following table:

\begin{table}[H]
\label{tab:slimgb}
	\begin{center}
	    \caption{(Dummy)Runtime of Gr\"obner Basis Computation of Mastrovito Multipliers in Singular using ATO $>$.}\label{tab:ATO}
	    \begin{tabular}{|c|c||c|} 
	        \hline
		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (minutes)   \\
		\hline
	        $16$	&  $1,871$  & $2.4$ \\
		$24$	&  $3,135$  & $12$  \\
	        $32$	&  $5,549$  & $22.6$ \\
	        $40$	&  $8,587$  & $266$ \\
		$48$	& $12,327$  & NA (Out of Memory) \\
	        \hline
	    \end{tabular}
	\end{center} 
\end{table}

It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 40(dummy).
Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} is also available to 
accelerate the GB computation here. More specifically, we obviate the GB computation by turning the Buchberger's algorithm 
into a single-step multivariate polynomial division.

\begin{itemize}
\item First, a set of polynomials is constructed by polynomials translated from each logic gates, and they generate 
ideal $J_{gates}$. Similarly, ideal $J_{W-B}$ is generated by polynomials representing the word-bit correspondences.
They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{W-B}$.
\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using reverse topological order
in circuit structure, followed by word-level NS output (e.g. $R'$ in Figure \ref{fig:sequential}), 
then word-level PS inputs (e.g. $A,B,R$ in Figure \ref{fig:sequential}).
\item When computing the Gr\"obner basis of $J_{ckt}+J_0$ (adding vanishing polynomials), because of 
RATO, there only exist one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
\item Then we only need to compute the S-poly for $(f_w,f_g)$, the reduce the S-poly by $J_{ckt}+J_0$.
\end{itemize}

@@

\begin{Example}
\label{ex:newRATO}
From the circuit topological structure in Figure \ref{fig:5bitRH}, the term order under RATO is:
\begin{align}
&\{r_0',r_1',r_2',r_3',r_4'\}>\{r_0,r_1,r_2,r_3,r_4\}\nonumber\\
&>\{e_0,e_3,e_4\},\{d_0,d_1,d_2\},\{c_1,c_2,c_3,c_4\}\nonumber\\
&>\{a_0,a_1,a_2,a_3,a_4,b_0,b_1,b_2,b_3,b_4\}>R'>R>\{A,B\}\nonumber
\end{align}
The variables in braces are at the same topological level, they are arranged in LEX order. 
We search among all generators of $J_{ckt}$ from Example \ref{ex:RHSMPO} using RATO, 
and find a pair of polynomials whose leading 
monomials are not relatively prime:
$(f_w,f_g), f_w = r_0'+r_4+e_0, f_g =r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18} + R'$.
We calculate $Spoly$ can reduce it by $J_{ckt}+J_0$:
\begin{align}
&Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}\nonumber\\
&(\alpha^3+\alpha^2+\alpha) r_1+(\alpha^4+\alpha^3+\alpha^2) r_2+(\alpha^2+\alpha) r_3+(\alpha) r_4\nonumber\\
&+(\alpha^3+\alpha^2) a_1 b_1+(\alpha^4+\alpha^3+\alpha^2+\alpha) a_1 b_2+(\alpha^2+\alpha) a_1 b_3\nonumber\\
&+(\alpha^2+1) a_1 b_4+(\alpha^4+1) a_1 B+(\alpha^4+\alpha) a_2 b_1+(\alpha^4+\alpha^3+\alpha) a_2 b_2\nonumber\\
&+(\alpha^3+1) a_2 b_3+(\alpha^3+\alpha^2+1) a_2 b_4+(\alpha^3+\alpha^2) a_2 B+(\alpha^2+\alpha) a_3 b_1\nonumber\\
&+(\alpha^3+1) a_3 b_2+(\alpha+1) a_3 b_3+(\alpha^4+\alpha^2+\alpha) a_3 b_4\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) a_3 B+(\alpha^3+1) a_4 b_1+a_4 b_2+(\alpha^4+\alpha^2+\alpha) a_4 b_3\nonumber\\
&+(\alpha^4+\alpha^3+1) a_4 b_4+(\alpha^2+\alpha) a_4 B+(\alpha^4+1) b_1 A+(\alpha^3+\alpha^2) b_2 A\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) b_3 A+(\alpha^2+\alpha) b_4 A+(\alpha^4+\alpha^2+\alpha+1) R'+R+A B\nonumber
\end{align}
\end{Example}
Above example indicates that RATO based abstraction on 5-bit RH-SMPO will result a remainder containing both
bit-level variables and word-level variables. Moreover, the number of variables is still large such that
Gr\"obner basis computation will be inefficient.

>>>>

\chapter{Functional Verification of Sequential Normal Basis Multipliers}
\label{ch:normal}
With the size of sequential arithmetic circuits increasing rapidly, 
conventional methods become inefficient on verifying the correctness of these circuits.
In Chapter \ref{ch:reacha} we proposed a FSM traversal approach at word-level 
based on abstractions. Since the sequential arithmetic circuits can be modeled 
as FSMs, it is possible to apply our word-level traversal algorithm to solving 
the functional verification problem on these circuits.
From the work of Lv \cite{lv:phd},
we learn that arithmetic circuits in Galois field (GF) can be
represented using an ideal of polynomials, and the 
ideal generators themselves form a GB under a specialized term order
obtained by a reverse topological traversal of the circuit.
Furthermore, according to the work of Pruss {\it et al.}
\cite{pruss:tcad15}, we can also abstract the word-level signature of an arithmetic 
component working in GF. Thus, we consider exploiting the concepts of \cite{lv:phd} and \cite{pruss:tcad15}
to efficiently perform abstraction of the state-space to specific sequential GF circuits.
In each time-frame, we can use the techniques 
from \cite{pruss:tcad15} to abstract the word-level
signature of the combinational logic, which corresponds
to the transition function in our traversal algorithm.
We show how this approach can be applied to verify sequential GF multipliers
which we can apply our traversal algorithm to actually 
verify its functional correctness.

\section{Motivation}
\label{sec:normal_motiv}
% From the preliminaries about FSMs in Section \ref{sec:reach}, we learn that the
% Moore machine does not rely on inputs for state transitions. 
As depicted in Figure \ref{fig:Moore}(a), a typical Moore machine implementation
consists of combinational logic component and register files, where
$r_0,\dots,r_k$ are present state (PS) variables 
standing for pseudo inputs, and $r_0',\dots,r_k'$ are next state (NS) variables standing for
pseudo outputs. Figure \ref{fig:Moore}(b) shows the state transition graph (STG) of a
GF arithmetic circuit implementation of 
a Moore machine with $k+1$ distinct states. We notice that it forms a simple chain,
with $k$ consecutive transitions the machine reaches final state $R_k$.

@@

In practice,
some arithmetic components are designed in sequential circuits similar to the structure in 
Figure \ref{fig:Moore}(a). Initially the operands are loaded into the registers, 
then the closed circuit executes without taking any additional information from outside,
and stores the results in registers after $k$ clock cycles. Its behavior can be described using the 
STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, 
$R_{init}$ is the initial state, $R_1$ to $R_{k-1}$ are intermediate results stored as NS variables 
of current time-frame and PS variables for the next time-frame, and $R_k$ (or $R_{final}$) 
is the final result given by arithmetic circuits (and equals to the
arithmetic function when circuit is working correctly).
This kind of design results in 
reusing a smaller combinational logic component such that the area cost is greatly optimized.
An example of such designs is the sequential GF arithmetic multipliers described in Chapter \ref{ch:prelim_GF}.
However, it also brings difficulties in verifying the the circuit functions.

@@

Conventional methods to such a sequential circuit may consist of unrolling the circuit for 
$k$ time-frames, and performing an equivalence checking between the unrolled machine and
the specification function. However, the number of gates will grow fast when unrolling
on bit-level. Meanwhile the structural similarity based equivalence checking techniques 
will fail when the sequential circuit is highly customized and optimized from the naive specification 
function. As a result, conventional techniques are grossly inefficient for large circuits.
Therefore, a new method based on our proposed word-level FSM traversal technique is worthy of exploration.

\section{Formal Verification of Normal Basis Multipliers using Gr\"obner Bases}
The gate-level design of a NB multiplier can be generated using approaches introduced in 
Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
an approach similar to that in Chapter \ref{ch:reacha}. 
First we introduce the sketch of our approach using abstraction term order (ATO)
mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
the concept of ``RATO", which is previously used in Section \ref{sec:improve}.
% The following part borrows contents
% from my own conference paper \cite{myDATE}.
\subsection{Implicit Unrolling based on Abstraction with ATO}
\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
\caption{Architecture of a combinational GF multiplier}
\label{fig:comb_RAB}}
\end{figure}

The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
$intermediate~variables~>~R~>~A,B$ for the circuit in Figure \ref{fig:comb_RAB} to compute Gr\"obner basis, 
the function of the combinational logic component can be abstracted as 
$$R = \Func(A,B)$$
To verify the functional correctness of a combinational NB multiplier ({\it e.g.} Mastrovito multiplier
or Montgomery multiplier), the function given by abstraction will be computed as
$$R=A\cdot B$$
While in the sequential case, the function of combinational logic only fulfills a part of the multiplication.
For example, in the RH-SMPO introduced in Section \ref{sec:nbdesign}, 
the combinational logic actually implements  $F_{k-1}$ in Equation \ref{eqn:aux},
while computing all of $F_i$ requires $k$-cycle unrolling of the circuit.
Nevertheless, the abstraction still provides a word-level 
representation which can be more efficient for unrolling than bit-level expressions.
In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
explicit unrolling and avoid bit-blasting problem.

For 2-input sequential NB multipliers, abstraction is utilized to implement following algorithm:
%\vspace{-0.1in}
\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Circuit polynomial ideal $J_{ckt}$, vanishing ideal $J_0$, initial
   state ideal for output $R (=0)$, inputs $\mathcal{G}(A_{init}), \mathcal{H}(B_{init})$} 
\KwOut{Circuit function $R_{final}$ after $k$ cycles of $A_{init}, B_{init}$}
  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  \tcc{$A, B$ are word-level variables, solutions to polynomial equations 
  $\mathcal{G},\mathcal{H}$ denote the initial values of preloaded operands $A_{init},B_{init}$}
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J_{ckt} + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	\tcc{Using projections of varieties from abstraction theory, we get NS variables $R',A',B'$
	in terms of PS $A,B$}
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
	\tcc{By replacing NS variables with PS variables we push it to next time-frame}
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
\DecMargin{1em}

@@

Given a sequential GF circuit in Figure \ref{fig:sequential} over $\Fkk=\F_2[x]$, irreducible polynomial $p(\alpha) = 0$
defines the primitive element $\alpha$, $\beta = \alpha^t$ is the normal element.
The circuit includes word-level variables $A, B, R$
denoting the {\it present states (PS)} and $A', B', R'$ denoting the {\it next
  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
store the result, and $A, B\ (A', B')$ store input operands. For example,
for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
0$) are the initial values (operands) loaded into the registers,  and
$R = \Func(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
result after $k$-cycles. Our approach aims to find this polynomial
representation for $R$.

Each gate in the combinational logic is represented by a Boolean
polynomial. To 
this set of Boolean polynomials, we append the polynomials that define
the word-level to bit-level relations for PS and NS variables ($A =
\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
as ideal $J_{ckt} = \langle 
f_1, \dots, f_s \rangle \subset \F_2[x_1, \dots, x_d, R, R', A, A', B,
  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
  of the circuit. The ideal of vanishing polynomials 
$J_0 = \langle x_i^2-x_i,R_j^{2^k}-R_j\rangle,~\forall \text{ bit-level variables }x_i\text{ and word-level variables }R_j$, is also included. 
Then the implicit FSM unrolling problem is setup for abstraction.

The configurations of the flip-flops are the states of the
machine. Since the set of states is a finite set of points, we
can consider it as the variety of an ideal related to the circuit.
Moreover, since we are interested in
the function encoded by the state variables (over $k$-time
frames), we can project this variety on the word-level state
variables, starting from the initial state $A_{init}, B_{init}$.
Projection of varieties (geometry) corresponds to elimination ideals
(algebra), and can be analyzed via \Grobner bases as shown in Section \ref{sec:abstraction}. Therefore, we
employ a \Grobner basis computation with ATO: we use a {\it lex term
  order} with {\it bit-level variables} 
$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
allows to eliminate all the bit-level variables 
%(corresponding to the combinational logic and the state variables),
%so as to 
and derives a representation only in terms of words. 
Consequently, $k$-successive \Grobner basis computations implicitly
unroll the machine, and provide a word-level algebraic $k$-cycle
abstraction for $R'$ as $R' = \Func(A_{init}, B_{init})$.

@@

In the practice of sequential GF multiplier verification, the combinational 
logic actually implements function not only related to current operands $A$ and $B$,
but also involved with PS variable ({\it i.e.} temporary product) $R$. Which can be 
obtained using the abstraction:
$$R' = \Func(A,B,R)$$

@@

\begin{Example}[Functional verification of 5-bit RH-SMPO]
\label{ex:RHSMPO}
Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO, with normal element $\beta = \alpha^5$. The transition function for
operands $A,B$ perform cyclic shift, while the transition function for $R$ has to be computed through the Gr\"obner basis
abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
all gates in combinational logic block and definition of word-level variables.
$$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$

@@

According to line 5 of Algorithm \ref{alg:modified}, we merge $J_{ckt}$, $J_0$ and $from^i$, then compute its
Gr\"obner basis with abstraction term order. 
Concretely, in the first iteration $from^0$ contains three generators. The first one describes PS variable $R$ --
temporary (or initial, in first iteration) product, which equals to 0 according to the mechanism of sequential 
GF multiplier. The polynomial is written as 
$$R$$
The other two polynomials describes PS variable $A,B$ -- current multiplication operands, and we write them 
in parameters $A_{init}, B_{init}$. Since this is the first iteration, they have the same form with $f_{16}$ and $f_{17}$:
\begin{align*}
& A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18} \\
& B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}
\end{align*}
After computing the Gr\"obner basis of $J_{ckt}+J_0+from^0$ using ATO 
$$\text{all bit-level variables} > R > \{R',A_{init},B_{init}\}$$
there is a polynomial in the form of $R'+\Func(A_{init},B_{init})$,
which should be included in $to^{i+1}$. $to^{i+1}$ also excludes next state variable $A'$ and $B'$, instead we 
redefine $A_{init}$ and $B_{init}$ using next state bit-level variables $\{a_i', b_j'\}$. Next state Bit-level variables
$a_i' = a_{i-1\pmod k}, b_j' = b_{j-1\pmod k}$ according to definition of cyclic shift.

@@

\subsection{Overcome Computational Complexity using RATO}
We implemented Algorithm \ref{alg:modified} in \textsc{Singular}, with similar setup and environments as
with the experiments in Section \ref{sec:exp_reacha}. The result of verifying Agnew's SMPO is shown 
in following table:

\begin{table}[H]
\label{tab:slimgb}
	\begin{center}
	    \caption{Runtime of Gr\"obner Basis Computation of Agnew's SMPO in Singular using ATO $>$.}\label{tab:ATO}
	    \begin{tabular}{|c|c||c|} 
	        \hline
		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (seconds)   \\
		\hline
	        $5$	&  $65$  & $20$ \\
		$11$	&  $275$  & $>35,000$  \\
	        \hline
	    \end{tabular}
	\end{center} 
\end{table}

It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 11 bits operands.
Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} can also be applied to 
accelerate the GB computation here. More specifically, we circumvent the GB computation 
of Line 5 in Algorithm \ref{alg:modified} by turning the Buchberger's algorithm 
into a single-step of multivariate polynomial division.

\begin{itemize}
\item First, a set of polynomials is constructed from each logic gates in the circuit, and they generate 
ideal $J_{gates}$. Similarly, ideal $J_{WB}$ is generated by polynomials representing the word-to-bit correspondences.
They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{WB}$.
\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using a reverse topological order
derived from the circuit structure, followed by word-level NS outputs ({\it e.g.} $R'$ in Figure \ref{fig:sequential}), 
then word-level PS inputs ({\it e.g.} $A,B,R$ in Figure \ref{fig:sequential}).
\item When computing the Gr\"obner basis of $J_{ckt}+J_0$ (adding vanishing polynomials), because of 
RATO, there exists only one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
\item Then we only need to compute the $Spoly$ for $(f_w,f_g)$, and reduce the $Spoly$ by $J_{ckt}+J_0$.
\end{itemize}

@@

\begin{Example}
\label{ex:newRATO}
From the circuit topological structure in Figure \ref{fig:5bitRH}, the term order under RATO is:
\begin{align}
&\{r_0',r_1',r_2',r_3',r_4'\}>\{r_0,r_1,r_2,r_3,r_4\}\nonumber\\
&>\{e_0,e_3,e_4\},\{d_0,d_1,d_2\},\{c_1,c_2,c_3,c_4\}\nonumber\\
&>\{a_0,a_1,a_2,a_3,a_4,b_0,b_1,b_2,b_3,b_4\}>R'>R>\{A,B\}\nonumber
\end{align}
The variables in braces are at the same topological level, they are arranged in LEX order. 
We search among all generators of $J_{ckt}$ from Example \ref{ex:RHSMPO} using RATO, 
and find a pair of polynomials whose leading 
monomials are not relatively prime:
$(f_w,f_g), f_w = r_0'+r_4+e_0, f_g =r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18} + R'$.
We calculate $Spoly$ can reduce it by $J_{ckt}+J_0$:
\begin{align}
&S\text{-}poly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}\nonumber\\
&(\alpha^3+\alpha^2+\alpha) r_1+(\alpha^4+\alpha^3+\alpha^2) r_2+(\alpha^2+\alpha) r_3+(\alpha) r_4\nonumber\\
&+(\alpha^3+\alpha^2) a_1 b_1+(\alpha^4+\alpha^3+\alpha^2+\alpha) a_1 b_2+(\alpha^2+\alpha) a_1 b_3\nonumber\\
&+(\alpha^2+1) a_1 b_4+(\alpha^4+1) a_1 B+(\alpha^4+\alpha) a_2 b_1+(\alpha^4+\alpha^3+\alpha) a_2 b_2\nonumber\\
&+(\alpha^3+1) a_2 b_3+(\alpha^3+\alpha^2+1) a_2 b_4+(\alpha^3+\alpha^2) a_2 B+(\alpha^2+\alpha) a_3 b_1\nonumber\\
&+(\alpha^3+1) a_3 b_2+(\alpha+1) a_3 b_3+(\alpha^4+\alpha^2+\alpha) a_3 b_4\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) a_3 B+(\alpha^3+1) a_4 b_1+a_4 b_2+(\alpha^4+\alpha^2+\alpha) a_4 b_3\nonumber\\
&+(\alpha^4+\alpha^3+1) a_4 b_4+(\alpha^2+\alpha) a_4 B+(\alpha^4+1) b_1 A+(\alpha^3+\alpha^2) b_2 A\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) b_3 A+(\alpha^2+\alpha) b_4 A+(\alpha^4+\alpha^2+\alpha+1) R'+R+A B\nonumber
\end{align}
\end{Example}
The above example indicates that RATO based abstraction on 5-bit RH-SMPO will result in a remainder containing both
bit-level variables and word-level variables. Moreover, the number of variables is still large such that
Gr\"obner basis computation will be inefficient.

Since the remainder from \emph{Spoly} reduction contains some bit-level variables, our objective is to compute
a polynomial that contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
this problem is to replace the bit-level variables by equivalent polynomials that only contain word-level
variables, {\it e.g.} $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.

\begin{Example}
\label{ex:BLVS}
{\bf Objective}:\ In $\F_{2^5}$ with normal element $\beta = \alpha^5$, compute polynomial 
$a_i + \mathcal{G}_i(A)$ from $f_0 = a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A
 = g_0 + A$.
